Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-09-01 ba_resubmit/kundinger_resubmit.tex
Paper edits

One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12} = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.

@@

\[L(\hat{Z_j}, Z_j)=\begin{cases} 
	0  & \text{if } Z_j = \hat{Z_j}; \\
	\theta_R,  & \text{if } \hat{Z_j} = R; \\
	\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_1 + j ; \\
	\theta_{01},  & \text{if } Z_j = n_1 + j,\hat{Z_j} \leq n_1 ; \\
	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
on the linkage status, \(\theta_{10}\) is the loss from a false
non-match, \(\theta_{01}\) is the loss from a false match, and
\(\theta_{11}\) is the loss from the special case of a false match in
which the record has a true match other than the one estimated by the
model.

@@

Although sampling \(Z_j\) from a the full conditional provided in Section~\ref{gibbs_sampling} conceptually straightforward, it becomes computationally burdensome when $n_1$ is large. This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$. To speed up computation, we break this sampling step into two. First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link. Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + j\). Then,

>>>>

One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.

@@

\[L(\hat{Z_j}, Z_j)=\begin{cases} 
	0  & \text{if } Z_j = \hat{Z_j}; \\
	\theta_R,  & \text{if } \hat{Z_j} = R; \\
	\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_1 + j ; \\
	\theta_{01},  & \text{if } Z_j = n_1 + j,\hat{Z_j} \leq n_1 ; \\
	\theta_{11},  & \text{if } Z_j \leq n_1, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
on the linkage status, \(\theta_{10}\) is the loss from a false
non-match, \(\theta_{01}\) is the loss from a false match, and
\(\theta_{11}\) is the loss from the special case of a false match in
which the record has a true match other than the one estimated by the
model.

@@

Although sampling \(Z_j\) from a the full conditional provided in Section~\ref{gibbs_sampling} is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large. This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$. To speed up computation, we break this sampling step into two. First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link. Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + j\). Then,
